## Antibody Modelling Function for mice
## Create Fig 4-A, S-Fig-3


######### Load shared functions & required packages ############
```{r}
source("functions.R")

required_packages <- c('dplyr', 'deSolve', 'ggplot2','purrr')
install_and_load(required_packages)
```

########### read parameter ############
```{r}
pop <- read.csv("populationParameters_mice.txt",  row.names = 1)
```

######## Set time range ###########
```{r}
Tmin <- 0
Tmax <- 180 
step_size <- 0.005
times<-c(seq(Tmin,Tmax,step_size))
```

######### ODE model, sampling and sample-ODE functions: see functions.R ############
## ode_ab_pop(), sample_ab(), run_ODE_control() are defined in functions.R.

############## Introduce the covariate of DOSE for different mice groups ################
```{r}
df_cov <- data.frame(Vaccine = c( "0.05ug Pfizer", 
                                  "0.1ug Pfizer", 
                                  "0.2ug Pfizer", 
                                  "0.5ug Pfizer",
                                  "1ug Pfizer")
                 )

df_cov
```
############## Simulate the trajectory ####################
```{r}
n = 1000  # Can change to 100, or change stepsize to a larger value to save time

# Vaccine-group dose effects on k1 and k2 (used in the population trajectory below)
beta_k1_0_1ug <- pop["beta_k1_Vaccine_Group_0_1ug_Pfizer", "value"]
beta_k1_0_2ug <- pop["beta_k1_Vaccine_Group_0_2ug_Pfizer", "value"]
beta_k1_0_5ug <- pop["beta_k1_Vaccine_Group_0_5ug_Pfizer", "value"]
beta_k1_1ug   <- pop["beta_k1_Vaccine_Group_1ug_Pfizer",   "value"]
beta_k2_0_1ug <- pop["beta_k2_Vaccine_Group_0_1ug_Pfizer", "value"]
beta_k2_0_2ug <- pop["beta_k2_Vaccine_Group_0_2ug_Pfizer", "value"]
beta_k2_0_5ug <- pop["beta_k2_Vaccine_Group_0_5ug_Pfizer", "value"]
beta_k2_1ug   <- pop["beta_k2_Vaccine_Group_1ug_Pfizer",   "value"]

Fit <- list()
pb <- txtProgressBar(min = 0, max = 2, style = 3)
for (g in 1:5){
  vac <- df_cov$Vaccine[g]
  par <- c(k1=exp(log(pop["k1_pop", "value"])+(vac=='0.1ug Pfizer')*beta_k1_0_1ug+(vac=='0.2ug Pfizer')*beta_k1_0_2ug+(vac=='0.5ug Pfizer')*beta_k1_0_5ug+(vac=='1ug Pfizer')*beta_k1_1ug),
           k2=exp(log(pop["k2_pop", "value"])+(vac=='0.1ug Pfizer')*beta_k2_0_1ug+(vac=='0.2ug Pfizer')*beta_k2_0_2ug+(vac=='0.5ug Pfizer')*beta_k2_0_5ug+(vac=='1ug Pfizer')*beta_k2_1ug),
           A0=pop["A0_pop", "value"],
           C = pop["peak_pop", "value"])
  
  best_fit <- ode_ab_pop(par)
  
  pars <- sample_ab(pop, n, vac)
  total_VL <- run_ODE_control(pars)
  
  Min90  <- apply(total_VL,1,function(x){quantile(x,0.025,na.rm=T)})
  Max90  <- apply(total_VL,1,function(x){quantile(x,0.975,na.rm=T)})
  Min50  <- apply(total_VL,1,function(x){quantile(x,0.25,na.rm=T)})
  Max50  <- apply(total_VL,1,function(x){quantile(x,0.75,na.rm=T)})
  Mean50  <- apply(total_VL,1,function(x){quantile(x,0.5,na.rm=T)})
  index <- which(Mean50 < -2)[1]
  # Check if an index was found, and then replace the value
  if (!is.na(index)) {
    Mean50[index] <- -2
  }  
  
  Fit[[g]] <- cbind(best_fit,Min90,Max90,Min50,Max50,Mean50,vac)
  Fit[[g]] <- data.frame(Fit[[g]])
  colnames(Fit[[g]]) <- c("time","best_fit","Min95","Max95","Min50","Max50","Mean50","Group")
  setTxtProgressBar(pb, g)
}
close(pb)
combine_pop <- map_df(Fit, ~as.data.frame(.x))

```
##############################
####### Figure 4-A ###########
##############################

```{r}
# Create the plot
plot_pop1 <- ggplot(combine_pop, aes(x = time, y = best_fit)) +
  geom_ribbon(aes(ymin = Min95, ymax = Max95, group = Group, fill = Group), alpha = 0.1) +
  geom_ribbon(aes(ymin = Min50, ymax = Max50,group = Group, fill = Group), alpha = 0.15) +
  geom_line(aes(group = Group, color = Group),linewidth = 1.5) +
  labs(
    #title = "Visualization of Antibody Response",
   x = "days since first vaccination",
       y = "anti-spike IgG (%binding)",
    color = "Group",
    fill = "Group"
  ) +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+ theme(
 # plot.title = element_text(size = 16),        
  axis.title.x = element_text(size = 14),       
  axis.title.y = element_text(size = 14),       
  axis.text = element_text(size = 14) ,         
  legend.title = element_text(size = 14),  
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)
plot_pop1
```

###########################################
########Individual Trajectory #############
###########################################

#read original data
```{r}
original <- read.csv("demo_data_mice.csv")
#note: this data is not provided.
colnames(original)[colnames(original) == "Mice.Number"] <- "Code"
original
```
####### load Monolix results ########
```{r}
Est <- read.csv("estimatedIndividualParameters_mice.txt", sep = ",", comment.char = "", header = T)
Simulated <- read.csv("simulatedIndividualParameters_mice.txt", sep = ",", comment.char = "", header = T)
max(Simulated$rep) #See how many simulated parameters for each individual
```
######## Reset time grid for individual fits & compute them ########
## ind_fit_mice() is defined in functions.R
```{r}
Tmin <- 0
Tmax <- 180
step_size <- 0.01
times<-c(seq(Tmin,Tmax,step_size))
n <- 100

individual_fit <- ind_fit_mice(Est, Simulated, original)
```
#########################################
########Supplementary Figure 3 ##########
#########################################

```{r}
plot_indfit <- ggplot(individual_fit, aes(x = Days)) +
  # Add shaded area for Min/Max range
  geom_ribbon(aes(ymin = Min, ymax = Max, fill = Vaccine.Group), alpha = 0.3) +
  # Add the fit line
  geom_line(aes(y = y, color = Vaccine.Group), size = 1, alpha = 0.8) +
  # Add individual's point
  geom_point(data = original, aes(x = Days, y = Antibody.Level, color = Vaccine.Group), 
             shape = 16, size = 2, alpha = 2) +
  # Separate plots for each individual
  facet_wrap(~ as.factor(Code)) +
  # Add labels and theme
  labs(
    #title = "Individual Fit with Shaded Uncertainty",
     x = "days since first vaccination",
       y = "anti-spike IgG (%binding)",
    fill = "Vaccine Group",
    color = "Vaccine Group"
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) + theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 12),        # X-axis label font size
  axis.title.y = element_text(size = 12),        # Y-axis label font size
  axis.text = element_text(size = 12) ,           # Axis tick labels font size
  legend.title = element_text(size = 12),  # Legend title font size
  legend.text = element_text(size = 10),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)

plot_indfit
```




