############################################################################################################
############## Mice antibody model paramter (growth rate, decay rate) associated with dose amount ##########
############################################################################################################

############# S-Fig-2 ###############
############# Fig-4-B ###############

```{r}
source("functions.R")

library(dplyr)
library(ggplot2)
library(deSolve)
library(stringr)
```
#############################################
########### Load Mice parameter #############
#############################################

```{r}
pop_data <- read.csv("populationParameters_mice.txt",  row.names = 1)

# Extract k1 and k2 base population values
k1_pop <- pop_data["k1_pop", "value"]
k2_pop <- pop_data["k2_pop", "value"]

# Extract covariates for k1 and k2
beta_k1 <- pop_data[grep("^beta_k1", rownames(pop_data)), ]
beta_k2 <- pop_data[grep("^beta_k2", rownames(pop_data)), ]

# Calculate k1 for each dose group
k1_values <- data.frame(
  Dose_Group = c('0_05ug_Pfizer', gsub("beta_k1_Vaccine_Group_", "", rownames(beta_k1))),
  k1 = c(k1_pop, exp(log(k1_pop) + beta_k1$value))
)

# Calculate k2 for each dose group
k2_values <- data.frame(
  Dose_Group = c('0_05ug_Pfizer',gsub("beta_k2_Vaccine_Group_", "", rownames(beta_k2))),
  k2 = c(k2_pop, exp(log(k2_pop) + beta_k2$value))
)

# Combine k1 and k2 results
result <- merge(k1_values, k2_values, by = "Dose_Group", all = TRUE)

# Print the result
print(result)
```

```{r}
# Create the 'dose' column by extracting the numeric dose from 'Vaccine.Group'
result <- result %>%
  mutate(dose = as.numeric(str_extract(str_replace(Dose_Group, "_", "."), "\\d+\\.?\\d*")))
result
```

###############################################################
############## TRY DIFFERENT REGRESSION ########################
###############################################################


##########Decay Rate K2###############
## r2_k2() is defined in functions.R

```{r}
exp_model <- nls(k2 ~ a*exp(-b*dose),data = result,start = list(a = 1, b = 1))
powerlaw_model <- nls(k2 ~  a * dose^(-b)  ,data = result, start = list(a = 1, b = 1))
linear_k2 <- lm(k2 ~ dose, data = result)
log_model_k2 <- nls(k2 ~ -(log(a*dose+b)),data = result,start = list(a =1, b =0.01))

# Model summary
summary(exp_model)
summary(powerlaw_model)
summary(linear_k2)
summary(log_model_k2)
print(paste("AIC linear model:",AIC(linear_k2)))
print(paste("AIC exp model:",AIC(exp_model)))
print(paste("AIC powerlaw model:",AIC(powerlaw_model)))
print(paste("AIC log model:",AIC(log_model_k2)))

print(paste("R^2 of exp model",r2_k2(exp_model, result)))
print(paste("R^2 of powerlaw model",r2_k2(powerlaw_model, result)))
print(paste("R^2 of linear model",r2_k2(linear_k2, result)))
print(paste("R^2 of logarithm model",r2_k2(log_model_k2, result)))
```


##########Growth Rate K1###############
## r2_k1() is defined in functions.R
```{r}
exp_model_k1 <- nls(k1 ~ a*exp(-b*dose),data = result,start = list(a = 1, b = 1))
linear_k1 <- lm(k1 ~ dose, data = result)
powerlaw_model_k1 <- nls(k1 ~  a * dose^(-b)  ,data = result, start = list(a = 1, b = 1))
log_model_k1 <- nls(k1 ~ (log(a*dose-b)),data = result,start = list(a =1, b =0.01),trace = TRUE)

summary(exp_model_k1)
summary(linear_k1)
summary(powerlaw_model_k1)
summary(log_model_k1)


print(paste("AIC linear model:",AIC(linear_k1)))
print(paste("AIC exp model:",AIC(exp_model_k1)))
print(paste("AIC powerlaw model:",AIC(powerlaw_model_k1)))
print(paste("AIC log model:",AIC(log_model_k1)))

print(paste("R^2 of exp model",r2_k1(exp_model_k1, result)))
print(paste("R^2 of powerlaw model",r2_k1(powerlaw_model_k1, result)))
print(paste("R^2 of linear model",r2_k1(linear_k1, result)))
print(paste("R^2 of logarithm model",r2_k1(log_model_k1, result)))
```
####################################################
############k1-dose:log; k2-dose:powerlaw###########
####################################################

#####################################
############ Plot S-Fig-2############
#####################################

########## S-Fig-2-A #########
```{r}
# Calculate the CI for any regression, knowing the estimation and s.e.
# log regression for k1
set.seed(123)

# Parameters from the regression
a_est <- 0.36989    
a_se <- 0.02605
b_est <- 1.46569
b_se <- 0.01186

# Number of simulations to calculate CI
n_sim <- 10000

# Simulate parameter values from normal distributions
a_sim <- rnorm(n_sim, mean = a_est, sd = a_se)
b_sim <- rnorm(n_sim, mean = b_est, sd = b_se)

# Dose range
dose <- seq(0.001, 1.2, by = 0.001)

# Matrix to store predictions
k1_sim <- matrix(NA, nrow = n_sim, ncol = length(dose))

# Calculate k1 for each simulated parameter pair across dose range
for (i in 1:n_sim) {
  k1_sim[i, ] <- log(a_sim[i] * dose+ b_sim[i])
}

# Calculate 95% CI (2.5th and 97.5th percentiles) for each dose
ci_lower <- apply(k1_sim, 2, quantile, probs = 0.025) # 2 means apply on rows, this k2_sims has row number < no.simmulation, col number < no.dose range
ci_upper <- apply(k1_sim, 2, quantile, probs = 0.975)
k1_pred <- log(a_est * dose + b_est)  # Point estimate

# Combine results into a data frame
results <- data.frame(dose = dose, k1_pred = k1_pred, ci_lower = ci_lower, ci_upper = ci_upper)

#####################################################
head(results)

```

```{r}
plot1 <- ggplot(data = results, aes(x = dose)) +
  
  geom_line(aes(y = k1_pred, color = 'Fitted Line (powerlaw)'), size = 2) +
  
  geom_point(data = result, aes(x = dose, y = k1, color = 'Original Data'), size = 3) +
  
  geom_ribbon(data = results, aes(x = dose, ymin = ci_lower, ymax = ci_upper, color = '95% Confidence Interval'), alpha = 0.4, fill = "pink") +
  labs(color = 'Legend', 
       size = 'Point Size',
        x = 'Dose',
       y = ' k1') +
  
  scale_color_manual(values = c('Original Data' = 'red', 'Fitted Line (powerlaw)' = 'grey', '95% Confidence Interval' = 'pink')) +

   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='none',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+
  theme(
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "cm")
)
plot1
```
########## S-Fig-2-B #########

```{r}
# Calculate the CI for any regression, knowing the estimation and s.e.
# power law 
set.seed(123)

# Parameters from the regression
a_est <- 2.858e-04
a_se <- 9.654e-05
b_est <- 1.094
b_se <- 1.190e-01

# Number of simulations to calculate CI
n_sim <- 10000

# Simulate parameter values from normal distributions
a_sim <- rnorm(n_sim, mean = a_est, sd = a_se)
b_sim <- rnorm(n_sim, mean = b_est, sd = b_se)

# Dose range
dose <- seq(0.001, 1.2, by = 0.001)

# Matrix to store predictions
k2_sim <- matrix(NA, nrow = n_sim, ncol = length(dose))

# Calculate k2 for each simulated parameter pair across dose range
for (i in 1:n_sim) {
  k2_sim[i, ] <- a_sim[i] * dose^(-b_sim[i])
}

# Calculate 95% CI (2.5th and 97.5th percentiles) for each dose
ci_lower <- apply(k2_sim, 2, quantile, probs = 0.025) # 2 means apply on rows, this k2_sims has row number < no.simmulation, col number < no.dose range
ci_upper <- apply(k2_sim, 2, quantile, probs = 0.975)
k2_pred <- a_est * dose^(-b_est)  # Point estimate

# Combine results into a data frame
results <- data.frame(dose = dose, k2_pred = k2_pred, ci_lower = ci_lower, ci_upper = ci_upper)

#####################################################
head(results)
```
```{r}
plot2 <- ggplot(data = results, aes(x = dose)) +
  
  # Geom line with customized legend label
  geom_line(aes(y = k2_pred, color = 'Fitted Line (powerlaw)'), size = 2) +
  
  # Geom points with customized legend label and fixed size
  geom_point(data = result, aes(x = dose, y = k2, color = 'Original Data'), size = 3) +
  
  geom_ribbon(data = results, aes(x = dose, ymin = ci_lower, ymax = ci_upper, color = '95% Confidence Interval'), alpha = 0.4, fill = "pink") +
  # Add title and customize the legends for color and size
  labs(#title = 'Decay Rate k2 versus Dose (powerlaw)', 
       color = 'Legend', 
        x = 'Dose (μg)',
       y = ' k2') +
  
  # Customize color legend manually
  scale_color_manual(values = c('Original Data' = 'red', 'Fitted Line (powerlaw)' = 'grey', '95% Confidence Interval' = 'pink')) +

   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='none',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+
  coord_cartesian(ylim = c(0, 0.02)) + theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "cm")
)
  
plot2
```
####################################################################################################
####################################################################################################

######## Extrapolate k1 and k2 ##############

```{r}
new_doses <- data.frame(dose = seq(0.001, 1, by = 0.001))
head(new_doses)

predictions <- predict(log_model_k1, newdata = new_doses)
new_doses$k1 <- predictions
predictions2 <- predict(powerlaw_model, newdata = new_doses)
new_doses$k2 <- predictions2

new_doses_clean <- new_doses %>%
  dplyr::select(dose, k2, k1) 
new_doses_clean
#write.csv(new_doses_clean,"full_mice_dose_pars.csv", row.names = FALSE)
```
```{r}
target_doses <- seq(0.050, 1, by = 0.050)

# Filter the rows where the dose matches the target doses
filtered_doses <- new_doses %>%
  filter(round(dose, 3) %in% round(target_doses, 3)) %>%
  dplyr::select(dose, k1, k2) %>%

# View the filtered data
print(filtered_doses)
```

########## Allometric Scaling ################
```{r}
human_dose <- data.frame(dose = seq(5, 100, by = 5),
                         dose_mice = seq(0.05,1,by =0.05),
                         
                         k1 = NA,
                         k2 = NA)
human_dose
```

#######Human Parameter: dose = 30ug #######
```{r}
human_result <- read.csv("populationParameters_human.txt", row.names = 1)
k1_pop_human <- human_result["k1_pop", ]
k2_pop_human <- human_result["k2_pop", ]
human_parameter <- data.frame(dose = 30,
                              k1=k1_pop_human$value,
                              k2=k2_pop_human$value,
                              k1_lwr =k1_pop_human$P2.5_sa,
                              k1_upr =k1_pop_human$P97.5_sa,
                              k2_lwr =k2_pop_human$P2.5_sa,
                              k2_upr =k2_pop_human$P97.5_sa)
human_parameter
```
######Perform the scaling with SF=100###############

```{r}
scaling_result <- allometric_scaling(filtered_doses, 100, human_parameter)
scaling_result
```
```{r}
dose_dependent <- scaling_result
pop <- read.csv("populationParameters_human.txt", row.names = 1)
c = pop["peak_pop", "value"]
c
```


#####ODE function (ode_ab_pop) is defined in functions.R#######

########Run ODE for extrapolated Parameters############
```{r}
human_doses <- seq(10, 100, 5)
times <- seq(0,180,0.01)
Tmin <- 0
Tmax <- 180 
step_size <- 0.001
times<-c(seq(Tmin,Tmax,step_size))
Fit <- list()

final_df <- data.frame(Days = numeric(0), y = numeric(0), dose = numeric(0))
# Loop through each dose in human_doses
for (i in round(seq_along(human_doses)) ){
  
   dose <- human_doses[i]
  # Match the corresponding mice dose (divide by 100)
  mice_dose <- dose / 100
  
  # Extract k1 and k2 values for this dose
  par <- c(
    k1 = dose_dependent[round(dose_dependent$dose_human) == dose, "k1_human"],
    k2 = dose_dependent[round(dose_dependent$dose_human) == dose, "k2_human"],
    A0 = pop["A0_pop", "value"],
    C  = pop["peak_pop", "value"]
  )
  
  best_fit <- ode_ab_pop(par)
  
  # Create a dataframe for this dose with Days, y, and the current dose
  d1 <- data.frame(Days = times, y = best_fit$A, dose = rep(dose, length(times)))
  final_df <- rbind(final_df, d1)
}
final_df
```

#########################################
#####Figure4-B###########################
#########################################
```{r}
# Filter data
filtered_data <- final_df %>%
  filter(dose %in% seq(10, 100, by = 5))
 
original_data <- final_df %>%
  filter(dose == 30)  # Original data for dose = 30

dose_colors <- scales::seq_gradient_pal("lightblue", "darkblue", "Lab")(seq(0, 1, length.out = 19))
names(dose_colors) <- as.character(c(seq(10, 25, by = 5),seq(35,100,5)))  # Assign dose values as names
all_colors <- c(dose_colors, "Original Data (30ug)" = "red")

plot_huaman_dose = ggplot() +
  geom_line(data = filtered_data, aes(x = Days, y = y, group = dose, color = as.factor(dose)),size = 0.8
            ) +
  # Original data in red
  geom_line(data = original_data, aes(x = Days, y = y, color = "Original Data (30ug)"), linewidth = 1.5) +
  # Apply the custom color scale
  scale_color_manual(name = "Dose (ug)", values = all_colors) +
  labs(
   # title = "Predicted Human Anti-Spike IgG Trajectories Across Doses Using Allometric Scaling",
       x = "days since first vaccination",
       y = "anti-spike IgG (%binding)") +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        #legend.position='none',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"))
plot_huaman_dose
```



